---
title: "DHARMa Practice"
author: "Margaret Janiczek"
date: "2023-10-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this document I play around with the   `DHARMa` package, using some code snippets from the [package tutorial](https://cran.r-project.org/web/packages/DHARMa/vignettes/DHARMa.html) as well as adding some of my own. 

# Example of "good" model

Here in the simulated data the outcome is generated from a Poisson distribution. 
We fit a GLM with logit link function and use `DHARMa` to evaluate the results- which do not indicate anything troubling with the model.

```{r message=FALSE, warning = FALSE}
#install.packages("DHARMa")
library(DHARMa)
library(lme4)
library(gtsummary)
library(MASS)
library(tidyverse)

set.seed(1219)
testData = createData(sampleSize = 250)

```

Let's peak at the simulated data: 

```{r}

hist(testData$Environment1)
hist(testData$observedResponse)

testData %>%
  select(observedResponse, Environment1, group) %>%
  tbl_summary(by = group)

fittedModel <- glmer(observedResponse ~ Environment1 + (1|group) , 
                     family = "poisson", data = testData)

testDispersion(fittedModel)

simulationOutput <- simulateResiduals(fittedModel = fittedModel)

plot(simulationOutput)

plotResiduals(simulationOutput, form = testData$group)

testDispersion(simulationOutput)
```

# Example of overdispersion

Next we generate simulated data with overdispersion. When we fit a poisson model to this data and use the `DHARMa` functions lots of red flags come up. 

```{r}
testData2 = createData(sampleSize = 200, overdispersion = 1.5, family = poisson())
fittedModel2 <- glm(observedResponse ~  Environment1 , family = "poisson", data = testData2)

```

Let's peak at the simulated data: 

```{r}

hist(testData2$Environment1)
hist(testData2$observedResponse)

testData2 %>%
  select(observedResponse, Environment1, group) %>%
  tbl_summary(by = group)


simulationOutput2 <- simulateResiduals(fittedModel = fittedModel2)
plot(simulationOutput2)

testDispersion(simulationOutput2)
```


Finally I fit a negative binomial model to that overdispersed data, simulated new output and tested the results using `DHARMa`. Now it does not detect significant deviations!

```{r}

fittedModel_negbin<- glm.nb(observedResponse ~  Environment1 , data = testData2)
simulationOutput3 <- simulateResiduals(fittedModel = fittedModel_negbin)
plot(simulationOutput3)
testDispersion(simulationOutput3)
```

